### 2024-11-26 paper/paper.qmd
update model and paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(here)

# Load the cleaned speed data
data <- read_csv(here::here("data/02-analysis_data/analysis_speed_data.csv"))

# Ensure the data is ready for analysis
data <- data %>%
  mutate(
    month = as.Date(month)  # Ensure the 'month' column is in Date format
  )
```

# Introduction

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Traffic Speed Dataset

The traffic speed dataset contains information collected from multiple locations and months, detailing percentile speeds, total traffic volume, and counts of vehicles traveling within specific speed ranges. The dataset aggregates these metrics monthly, offering insights into traffic behavior and variations over time.

This dataset includes:
- Percentile speeds (5th, 10th, ... 95th) representing speed thresholds exceeded by corresponding proportions of vehicles.
- Traffic volume as the total number of vehicles observed monthly.
- Speed bins (e.g., 0-4 km/h, 5-9 km/h) for detailed breakdowns of vehicle counts across speed ranges.
- The table below, @tbl-speed-preview, showcases a preview of the cleaned dataset with selected columns to highlight key features.

```{r}
#| label: tbl-speed-preview
#| tbl-cap: Preview of cleaned traffic speed dataset.
#| echo: false
#| tbl-pos: H

data |> 
  select(month, pct_05, pct_50, pct_95, spd_30, spd_50, spd_100_and_above, volume) |> 
  head(10) |> 
  kable(
    col.names = c("Month", "5th Percentile", "Median (50th Percentile)", "95th Percentile", 
                  "Vehicles (30-34 km/h)", "Vehicles (50-54 km/h)", "Vehicles (>100 km/h)", "Total Volume"),
    booktabs = TRUE
  ) |> 
  column_spec(1, width = "2.5cm") |> 
  column_spec(2:8, width = "2.0cm")

```

## Random Sample from the Dataset
@tbl-random-sample provides a random sample of 10 observations, offering an unbiased snapshot of key metrics such as speed percentiles, total traffic volume, and counts of vehicles exceeding 100 km/h. This selection illustrates the dataset's diversity without emphasizing specific patterns.

```{r}
#| label: tbl-random-sample
#| tbl-cap: Random sample of traffic speed dataset.
#| echo: false
#| tbl-pos: H

set.seed(304)  # For reproducibility
data |> 
  sample_n(10) |> 
  select(month, pct_05, pct_50, pct_95, volume, spd_100_and_above) |> 
  kable(
    col.names = c("Month", "5th Percentile", "Median (50th Percentile)", "95th Percentile", "Total Volume", "Vehicles (>100 km/h)"),
    booktabs = TRUE
  ) |> 
  column_spec(1, width = "2.5cm") |> 
  column_spec(2:6, width = "2.0cm")

```

## Extreme Speeds Dataset
Vehicles traveling at extreme speeds (over 100 km/h) represent a critical factor in understanding traffic safety and violations. The table below, @tbl-extreme-speeds, highlights the top 5 months with the highest counts of vehicles exceeding 100 km/h. This provides insights into when extreme speeds are most prevalent.

```{r}
#| label: tbl-extreme-speeds
#| tbl-cap: Top 5 months with the highest counts of vehicles exceeding 100 km/h.
#| echo: false
#| tbl-pos: H

data |> 
  arrange(desc(spd_100_and_above)) |> 
  head(5) |> 
  select(month, volume, spd_100_and_above, pct_95) |> 
  kable(
    col.names = c("Month", "Total Volume", "Vehicles (>100 km/h)", "95th Percentile Speed"),
    booktabs = TRUE
  ) |> 
  column_spec(1, width = "2.5cm") |> 
  column_spec(2:4, width = "2.0cm")

```

## Proportion of Vehicles at Moderate Speeds
Another key insight from this dataset is the proportion of vehicles traveling within a moderate speed range (50–70 km/h). This measure helps understand how typical driving speeds align with safe and efficient traffic flow.

The table below, @tbl-speed-proportion, highlights monthly proportions of vehicles in this speed range.

```{r}
#| label: tbl-speed-proportion
#| tbl-cap: Monthly proportions of vehicles traveling at moderate speeds (50–70 km/h).
#| echo: false
#| tbl-pos: H

moderate_speed <- data %>%
  mutate(month = lubridate::month(month, label = TRUE, abbr = TRUE)) %>%
  group_by(month) %>%
  summarise(
    moderate_speed_count = sum(spd_50 + spd_55 + spd_60 + spd_65 + spd_70, na.rm = TRUE),
    total_volume = sum(volume, na.rm = TRUE),
    proportion_moderate = round((moderate_speed_count / total_volume) * 100, 2)
  )

moderate_speed |> 
  kable(
    col.names = c("Month", "Moderate Speed Vehicles", "Total Volume", "Proportion (%)"), 
    booktabs = TRUE
  ) |> 
  column_spec(1, width = "2.5cm") |> 
  column_spec(2:4, width = "2.5cm")

```

```{r}
#| label: fig-extreme-speeds-vs-volume
#| fig-cap: Relationship between extreme speeds (over 100 km/h) and total traffic volume, categorized by pre-2020 and post-2020 time periods.
#| echo: false
#| warning: false
#| message: false

# Process the data for analysis
extreme_vs_volume <- data %>%
  group_by(month) %>%
  summarise(
    extreme_count = sum(spd_100_and_above, na.rm = TRUE),
    total_volume = sum(volume, na.rm = TRUE)
  ) %>%
  mutate(period = ifelse(month < as.Date("2020-01-01"), "Pre-2020", "Post-2020"))

# Plot Extreme Speeds vs Total Traffic Volume by Time Period
ggplot(extreme_vs_volume, aes(x = total_volume, y = extreme_count, color = period)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Total Traffic Volume",
    y = "Extreme Speeds (Over 100 km/h)",
    color = "Time Period"
  ) +
  theme_minimal()

```

```{r}
#| label: fig-key-percentiles
#| fig-cap: Trends in key speed percentiles (5th, 50th, and 95th) and the spread between the 95th and 5th percentiles over time.
#| echo: false
#| warning: false
#| message: false
# Summarize Key Percentiles and Calculate Speed Spread
key_percentiles <- data %>%
  group_by(month) %>%
  summarise(
    pct_05 = mean(pct_05, na.rm = TRUE),
    pct_50 = mean(pct_50, na.rm = TRUE),
    pct_95 = mean(pct_95, na.rm = TRUE)
  ) %>%
  mutate(speed_spread = pct_95 - pct_05)  # Calculate spread

# Plot Key Percentiles and Spread
ggplot(key_percentiles, aes(x = month)) +
  geom_line(aes(y = pct_05, color = "5th Percentile")) +
  geom_line(aes(y = pct_50, color = "Median (50th Percentile)")) +
  geom_line(aes(y = pct_95, color = "95th Percentile")) +
  geom_line(aes(y = speed_spread, color = "Speed Spread"), linetype = "dashed") +
  labs(
    x = "Month",
    y = "Speed (km/h)",
    color = "Legend"
  ) +
  theme_minimal()

```

```{r}

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -1,69 +1,39 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Purpose: Tests the structure and validity of the analysis speed dataset
+# Author: Andy Jiang
+# Date: 25 November 2024
+# Contact: ayx.jiang@mail.utoronto.ca
+# Pre-requisites: Cleaned data generated and saved
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
+data <- read_csv("../data/02-analysis_data/analysis_speed_data.csv")
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+
+# Test that the dataset has 32000 rows (as expected from raw data summary)
+test_that("dataset has 32000 rows", {
+  expect_equal(nrow(data), 32000)
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+# Test that the dataset has 44 columns (since '_id' was removed from 45 original columns)
+test_that("dataset has 44 columns", {
+  expect_equal(ncol(data), 44)
 })
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
+# Test that the `month` column contains valid dates
+test_that("'month' column contains valid dates", {
+  expect_true(all(!is.na(as.Date(data$month, format = "%Y-%m-%d"))))
 })
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
+# Test that all percentile columns (`pct_05` to `pct_95`) are in increasing order
+percentile_columns <- data %>%
+  select(starts_with("pct_"))
+
+test_that("percentiles are in increasing order", {
+  expect_true(all(apply(percentile_columns, 1, function(row) all(diff(row) >= 0))))
 })
 
[…]

### 2024-11-26 scripts/05-exploratory_data_analysis.R

@@ -1,37 +1,87 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Perform exploratory data analysis (EDA) on the speed dataset
+# Author: [Your Name]
+# Date: [Current Date]
+# Contact: [Your Email]
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Ensure the dataset has been cleaned and saved in the 
+# appropriate location before running this script.
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Load Data ####
+# Read the cleaned data
+analysis_data <- read_csv("data/02-analysis_data/analysis_speed_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+#### Summary Statistics ####
+# Basic summary of the dataset
+summary(analysis_data)
+
+# Summary of speed percentiles (pct_05 to pct_95)
+speed_summary <- analysis_data %>%
+  select(starts_with("pct_")) %>%
+  summarise(across(everything(), list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE))
+
+print("Summary of speed percentiles:")
+print(speed_summary)
+
+#### Visualize Relationships ####
+
+# Distribution of speed percentiles
+analysis_data %>%
+  select(starts_with("pct_")) %>%
+  pivot_longer(cols = everything(), names_to = "percentile", values_to = "value") %>%
+  ggplot(aes(x = value, fill = percentile)) +
+  geom_density(alpha = 0.6) +
+  labs(
+    title = "Distribution of Speed Percentiles",
[…]

### 2024-11-26 scripts/06-model.R

@@ -1,0 +1,60 @@
+# Load libraries
+library(rstanarm)
+library(tidyverse)
+
+# Load the data
+analysis_data <- read_csv("data/02-analysis_data/analysis_speed_data.csv")
+
+# Step 1: Create a new response variable (spd_mean) if not already present
+# If 'spd_mean' does not exist, calculate it
+analysis_data <- analysis_data %>%
+  mutate(
+    spd_mean = rowMeans(select(., starts_with("pct_")), na.rm = TRUE)
+  )
+
+# Step 2: Filter data to remove rows with missing values
+filtered_data <- analysis_data %>%
+  filter(
+    !is.na(spd_mean),  # Ensure the response variable has no missing values
+    !is.na(volume),    # Ensure predictors have no missing values
+    !is.na(pct_05),    # Include pct_05
+    !is.na(pct_50),
+    !is.na(pct_95)
+  )
+
+# Step 3: Fit the Bayesian linear regression model
+mean_speed_model <- stan_glm(
+  formula = spd_mean ~ volume + pct_05 + pct_50 + pct_95,
+  data = filtered_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 123,
+  iter = 2000,
+  chains = 4,
+  control = list(adapt_delta = 0.9)  # Adjust for convergence
+)
+
+# Step 4: Save the model
+saveRDS(
+  mean_speed_model,
+  file = "models/speed_model.rds"
+)
+
+# Step 5: Add predicted values to the filtered dataset
+filtered_data <- filtered_data %>%
+  mutate(
+    predicted = predict(mean_speed_model, newdata = filtered_data)
+  )
+
+# Step 6: Plot Actual vs Predicted Mean Speeds
+ggplot(filtered_data, aes(x = predicted, y = spd_mean)) +
+  geom_point(alpha = 0.5, color = "blue") +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(
+    title = "Actual vs Predicted Mean Speeds (Using `pct_05`, `pct_50`, and `pct_95`)",
+    x = "Predicted Mean Speeds",
+    y = "Actual Mean Speeds"
+  ) +
+  theme_minimal()

### 2024-11-26 paper/paper.qmd
update paper, add model

@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Andy Jiang
+thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/toronto-vehicles-speed-analysis)."
 date: today
@@
 library(here)
+library(kableExtra)
+library(rstanarm)
 
@@
   )
+
+# **Add the data preprocessing steps to create 'filtered_data'**
+
+# Prepare the data used in the model
+analysis_data <- data
+
+# Step 1: Create a new response variable (spd_mean) if not already present
+analysis_data <- analysis_data %>%
+  mutate(
+    spd_mean = rowMeans(select(., starts_with("pct_")), na.rm = TRUE)
+  )
+
+# Step 2: Filter data to remove rows with missing values
+filtered_data <- analysis_data %>%
+  filter(
+    !is.na(spd_mean),  # Ensure the response variable has no missing values
+    !is.na(volume),    # Ensure predictors have no missing values
+    !is.na(pct_05),    # Include pct_05
+    !is.na(pct_50),
+    !is.na(pct_95)
+  )
+
 ```
@@
     booktabs = TRUE
-  ) |> 
-  column_spec(1, width = "2.5cm") |> 
-  column_spec(2:8, width = "2.0cm")
+  )
 
@@
     booktabs = TRUE
-  ) |> 
-  column_spec(1, width = "2.5cm") |> 
-  column_spec(2:6, width = "2.0cm")
+  )
 
@@
     booktabs = TRUE
-  ) |> 
-  column_spec(1, width = "2.5cm") |> 
-  column_spec(2:4, width = "2.0cm")
+  )
 
@@
     booktabs = TRUE
-  ) |> 
-  column_spec(1, width = "2.5cm") |> 
-  column_spec(2:4, width = "2.5cm")
+  )
 
@@
 
-```{r}
-
-```
-
 Talk way more about it.
@@
 first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+  readRDS(file = here::here("models/speed_model.rds"))
 ```

### 2024-11-29 scripts/00-simulate_data.R
update scripts

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulates vehicle speed data for pre-analysis
+# Purpose: Simulate traffic speed data for pre-analysis and testing purposes.
 # Author: Andy Jiang
 # Date: 23 November 2024
 # Contact: ayx.jiang@mail.utoronto.ca

### 2024-11-29 scripts/02-download_data.R

@@ -1,9 +1,11 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from Open Data Toronto
+# Purpose: Download and save traffic speed data from Open Data Toronto.
 # Author: Andy Jiang
 # Date: 23 November 2024
 # Contact: ayx.jiang@mail.utoronto.ca
-# Pre-requisites: None
+# Pre-requisites:
+#   - Internet access is required to fetch data from the Open Data Toronto API.
+#   - The `opendatatoronto` and `dplyr` libraries should be installed.
 
 
 #### Workspace setup ####

### 2024-11-29 scripts/03-clean_data.R

@@ -1,9 +1,13 @@
 #### Preamble ####
-# Purpose: Clean the traffic speed raw data
+# Purpose: Clean the raw traffic speed dataset to prepare it for analysis.
+# This includes removing unnecessary columns and saving the cleaned data.
 # Author: Andy Jiang
 # Date: 23 November 2024
 # Contact: ayx.jiang@mail.utoronto.ca
-# Pre-requisites: Raw data downloaded
+# Pre-requisites: 
+#   - Raw traffic speed data should be downloaded and available at 
+#     "data/01-raw_data/raw_speed_data.csv".
+#   - The `tidyverse` package should be installed.
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-29 scripts/04-test_analysis_data.R

@@ -1,9 +1,14 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the analysis speed dataset
+# Purpose: Validates the structure and integrity of the analysis speed dataset.
+# This script includes tests to verify dataset dimensions, column contents, 
+# and expected properties of specific variables such as speed percentiles.
 # Author: Andy Jiang
 # Date: 25 November 2024
 # Contact: ayx.jiang@mail.utoronto.ca
-# Pre-requisites: Cleaned data generated and saved
+# Pre-requisites: 
+#   - The cleaned speed dataset should be available at 
+#     "../data/02-analysis_data/analysis_speed_data.csv".
+#   - The `testthat` and `tidyverse` packages should be installed.
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-29 scripts/05-exploratory_data_analysis.R

@@ -1,11 +1,16 @@
 #### Preamble ####
-# Purpose: Perform exploratory data analysis (EDA) on the speed dataset
-# Author: [Your Name]
-# Date: [Current Date]
-# Contact: [Your Email]
-# License: MIT
-# Pre-requisites: Ensure the dataset has been cleaned and saved in the 
-# appropriate location before running this script.
+# Purpose: Perform exploratory data analysis (EDA) on the speed dataset.
+# This script generates summary statistics, visualizations, and insights 
+# about key variables such as speed percentiles and volume. It also includes 
+# correlation analysis and missing value analysis.
+# Author: Andy Jiang
+# Date: 25 November 2024
+# Contact: ayx.jiang@mail.utoronto.ca
+# Pre-requisites:
+#   - The cleaned speed dataset must be available at 
+#     "data/02-analysis_data/analysis_speed_data.csv".
+#   - The GGally library is required for correlation matrix visualization.
+#   - Ensure the "data/02-analysis_data/" directory exists to save the summary tables.
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-11-29 scripts/06-model.R

@@ -1,3 +1,13 @@
+#### Preamble ####
+# Purpose: Fit a Bayesian linear regression model to analyze traffic speed data 
+#          and predict mean speed using key predictors.
+# Author: Andy Jiang
+# Date: 25 November 2024
+# Contact: ayx.jiang@mail.utoronto.ca
+# Pre-requisites: The cleaned traffic speed data must be available in 
+#   - "data/02-analysis_data/analysis_speed_data.csv".
+#   - Necessary R packages (`rstanarm`, `tidyverse`) must be installed.
+
 # Load libraries
 library(rstanarm)
 library(tidyverse)

### 2024-11-30 paper/paper.qmd
update paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Analyzing Toronto Traffic Speeds: Insights into Patterns and Predictive Modeling"
+subtitle: "Bayesian Linear Regression Reveals Key Determinants of Mean Speed and Post-Pandemic Trends"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates traffic speed data from Toronto collected between 2017 and 2024, focusing on trends in speed percentiles and their relationship to traffic volume.
+Using a Bayesian linear regression model, we predict mean vehicle speeds based on key predictors such as 5th, 50th, and 95th percentile speeds and total traffic volume.
+Results highlight significant changes in traffic patterns post-2020, with increased extreme speeds linked to higher traffic volumes.
+This study provides insights into urban mobility dynamics, offering a basis for traffic management strategies and policy-making."
 format: pdf
@@
 bibliography: references.bib
+toc: true
 ---
@@
 library(rstanarm)
+library(modelsummary)
+library(dplyr)
+library(lubridate)
 
@@
 
-# **Add the data preprocessing steps to create 'filtered_data'**
-
 # Prepare the data used in the model
@@
 
-# Step 1: Create a new response variable (spd_mean) if not already present
 analysis_data <- analysis_data %>%
@@
 
-# Step 2: Filter data to remove rows with missing values
 filtered_data <- analysis_data %>%
@@
 
+# Read model
+mean_speed_model <- readRDS(file = here::here("models/speed_model.rds"))
 ```
@@
 # Introduction
+Traffic speed analysis is a critical tool for understanding urban mobility, road safety, and congestion.
+In metropolitan areas like Toronto, shifting traffic patterns reflect changes in urban infrastructure, population growth, and behavioral adaptations to external events, such as the COVID-19 pandemic.
+The advent of technologies like radar-equipped speed signs has enabled the collection of detailed traffic data, providing new opportunities to analyze and predict traffic behavior at a granular level.
+
+This paper focuses on traffic speed data collected in Toronto's designated Safety Zones between 2017 and 2024.
+The dataset comprises monthly summaries of speed percentiles, vehicle counts across specific speed ranges, and total traffic volumes.
+Notably, the study explores trends in extreme speeds (vehicles exceeding 100 km/h) and evaluates how they have evolved before and after 2020, a period marked by rapid population growth and post-pandemic recovery.
+These analyses aim to address a key gap: understanding how percentile-based speed measures and traffic volume influence overall mean speeds.
+
+To investigate these relationships, we employ a Bayesian linear regression model to predict mean vehicle speed (`spd_mean`) based on predictors including 5th, 50th, and 95th percentile speeds and traffic volume.
+Bayesian modeling allows for the incorporation of uncertainty and prior information, making it well-suited for analyzing urban traffic data.
+
+Our findings reveal significant changes in traffic behavior post-2020, including higher extreme speed occurrences at increased traffic volumes, reflecting broader shifts in urban mobility.
+This study highlights the importance of percentile-based metrics in predicting mean speeds and contributes to ongoing efforts to enhance traffic safety and efficiency.
+The paper is structured as follows: the data section provides an overview of the dataset and key variables, the model section outlines the Bayesian regression framework, results are discussed in detail, and the discussion section reflects on the implications and limitations of the findings.
 
@@
 
+All data analysis was conducted using R [@citeR] with the support of the following packages: tidyverse [@tidyverse], here [@here], ggplot2 [@ggplot2], rstanarm [@rstanarm], kableExtra [@kableExtra], modelsummary [@modelsummary], dplyr [@dplyr], and lubridate [@lubridate].
+These packages provided a robust framework for data manipulation, visualization, Bayesian modeling, and reporting, facilitating reproducible and efficient analysis.
+
 ## Overview
+This paper analyzes traffic speed data collected in Toronto over the period 2017–2024.
+The dataset contains detailed information on percentile speeds, total traffic volume, and counts of vehicles traveling within specific speed ranges.
+These data provide an opportunity to study the evolution of traffic patterns in Toronto, especially in the context of changing demographics, urbanization, and post-pandemic recovery.
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+The dataset aggregates metrics monthly, offering insights into how traffic behavior has changed over time.
+This analysis is particularly focused on extreme speeds (vehicles exceeding 100 km/h) and their relationship with total traffic volume, as well as trends in key speed percentiles.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset analyzed in this paper originates from the Safety Zone Watch Your Speed Program (WYSP), a City of Toronto initiative designed to monitor and influence driver behavior in designated Safety Zones.
+Data are collected using radar-equipped speed display signs installed on hydro poles or streetlights, which measure the speeds of oncoming vehicles with an accuracy of ±1 km/h and display the speeds on LED screens to encourage compliance with speed limits.
+These measurements are aggregated into monthly summaries, including key metrics such as percentile speeds (e.g., 5th, 50th, 95th percentiles), vehicle counts within specific speed ranges, and total observed traffic volume.
+The dataset is curated and published by Open Data Toronto, which ensures its metadata quality, completeness, and accessibility.
 
-## Traffic Speed Dataset
+However, the dataset has certain limitations.
+The number of vehicles recorded by the speed signs is not equivalent to true traffic volume, as it may exclude vehicles that pass too quickly or fall outside the radar’s range.
+Additionally, the presence of the signs might influence driver behavior, leading to changes in speed as drivers approach them.
+Furthermore, the data is limited to Safety Zones where the speed signs have been installed, restricting its spatial coverage.
+Despite these limitations, the dataset is well-documented, complete, and up-to-date, providing valuable insights into traffic speed patterns across Toronto.
 
+## Speed Dataset Preview
 The traffic speed dataset contains information collected from multiple locations and months, detailing percentile speeds, total traffic volume, and counts of vehicles traveling within specific speed ranges.
@@
 - Speed bins (e.g., 0-4 km/h, 5-9 km/h) for detailed breakdowns of vehicle counts across speed ranges.
-- The table below, @tbl-speed-preview, showcases a preview of the cleaned dataset with selected columns to highlight key features.
+- @tbl-speed-preview showcases a preview of the cleaned dataset with selected columns to highlight key features.
 
@@
 Vehicles traveling at extreme speeds (over 100 km/h) represent a critical factor in understanding traffic safety and violations.
-The table below, @tbl-extreme-speeds, highlights the top 5 months with the highest counts of vehicles exceeding 100 km/h.
+@tbl-extreme-speeds highlights the top 5 months with the highest counts of vehicles exceeding 100 km/h.
 This provides insights into when extreme speeds are most prevalent.
@@
 
-The table below, @tbl-speed-proportion, highlights monthly proportions of vehicles in this speed range.
+@tbl-speed-proportion highlights monthly proportions of vehicles in this speed range.
 
@@
 ```
+
+## Extreme Speeds and Traffic Volume
+
+@fig-extreme-speeds-vs-volume illustrates the relationship between extreme speeds (vehicles traveling over 100 km/h) and total traffic volume, categorized by pre-2020 and post-2020 periods.
+The post-2020 data points show a significant increase in both traffic volume and extreme speed counts, with some months recording over 2,000 vehicles exceeding 100 km/h.
+This increase may reflect changes in traffic behavior following the COVID-19 pandemic and a rapid population influx into Toronto through immigration.
+In contrast, pre-2020 data points display lower traffic volumes and fewer extreme speed counts, rarely exceeding 2,000.
+The fitted lines for each period highlight this divergence, with a steeper slope for post-2020 data, suggesting that higher traffic volumes post-pandemic are associated with a broader range of extreme speed occurrences.
 
@@
 
+## Trends in Key Percentiles
+
+@fig-key-percentiles examines trends in key speed percentiles (5th, 50th, and 95th percentiles) from 2017 to 2024, along with the average speed spread (the difference between the 95th and 5th percentiles).
+The 95th percentile speed demonstrates a clear decline, falling below 50 km/h over the years.
+Similarly, the 50th percentile speed dropped significantly between 2017 and 2018, stabilizing at around 30 km/h with minor fluctuations.
+The 5th percentile speed remained steady until mid-2022, when it experienced a noticeable decline to 5 km/h, down from its previous 10 km/h range.
+Despite these shifts, the speed spread remains relatively consistent, ranging between 35 and 40 km/h, indicating that the overall variability in speeds has not significantly changed.
+These trends may reflect broader changes in traffic regulations, driver behavior, or urban infrastructure developments in Toronto.
+
 ```{r}
@@
   labs(
-    x = "Month",
+    x = "Year",
     y = "Speed (km/h)",
@@
 ```
-
-Talk way more about it.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The purpose of this paper is to analyze the relationship between traffic volume, speed percentiles, and the mean speed of vehicles observed in Toronto.
+By utilizing a Bayesian linear regression model, we aim to investigate how various speed percentiles and total traffic volume predict the average vehicle speed (`spd_mean`).
+This model enables us to quantify the contributions of different speed measures and traffic density to overall traffic dynamics.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The Bayesian multiple linear regression model used in this paper predicts the mean speed (spd_mean) as a function of total traffic volume (volume) and speed percentiles (pct_05, pct_50, and pct_95).
+The model is defined as follows:
 
@@
 y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\mu_i &= \beta_0 + \beta_1 \cdot \text{volume}_{i} + \beta_2 \cdot \text{pct\_05}_{i} + \beta_3 \cdot \text{pct\_50}_{i} + \beta_4 \cdot \text{pct\_95}_{i} \\
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
 \sigma &\sim \mbox{Exponential}(1)
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+In the above model:
+
+- $\mu_i$ is the predicted mean speed (spd_mean) for the $i$-th observation given the total traffic volume and speed percentiles.
+- $\beta_0$ is the coefficient for the intercept.
+- $\beta_1$ is the coefficient for the predicted change in the mean speed given a one-unit increase in traffic volume (volume).
+- $\beta_2$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 5th percentile speed (pct_05).
+- $\beta_3$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 50th percentile speed (pct_50).
+- $\beta_4$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 95th percentile speed (pct_95).
+- $\sigma$ is the standard deviation of the residuals, representing unexplained variability in the mean speed.
+
+We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm]. We use the default priors from `rstanarm`.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The above model was chosen to investigate the relationship between the average speed (`spd_mean`) of vehicles and key predictors, including total traffic volume (`volume`) and speed percentiles (`pct_05`, `pct_50`, and `pct_95`).
+These variables were selected because they provide a comprehensive view of traffic patterns and vehicle behavior across different speed thresholds.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The percentiles represent critical thresholds of speed distribution, with the 5th percentile (`pct_05`) capturing the slowest vehicles, the 50th percentile (`pct_50`) representing the median speed, and the 95th percentile (`pct_95`) reflecting the fastest vehicles.
+These thresholds are expected to influence the average speed significantly.
+Similarly, total traffic volume (`volume`) accounts for overall road congestion and is anticipated to impact the mean speed inversely, as higher volumes may reduce average speeds due to congestion.
+
+We employ a Bayesian linear regression model implemented using the `rstanarm` package [@rstanarm].
+This approach provides a probabilistic framework for parameter estimation and allows the incorporation of prior information.
+Default priors were used for the regression coefficients (`Normal(0, 2.5)`) and residual standard deviation (`Exponential(1)`), as they are appropriate for general regression problems without overly constraining the estimates.
+
+The linearity assumption is reasonable given the nature of traffic data, where incremental changes in predictors like percentiles and volume are expected to have proportional effects on the mean speed.
+However, this model does not capture potential nonlinearities or interactions, which could be explored in future work if evidence suggests their relevance.
+
+We hypothesize that:
+
+- Higher 5th, 50th, and 95th percentile speeds will correspond to higher average speeds.
+- Higher traffic volumes will be associated with lower average speeds due to congestion effects.
+This model's simplicity balances interpretability with analytical rigor, making it a suitable choice for understanding traffic patterns and their determinants.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+## Model Coefficients {#sec-modelcoefficients}
+The results of the Bayesian linear regression model estimating the relationship between mean traffic speed (spd_mean) and its key predictors are summarized in @tbl-modelsummary.
 
 ```{r}
+#| label: tbl-modelsummary
+#| tbl-cap: Model summary of the predicted impact of traffic volume and speed percentiles on mean speed.
 #| echo: false
-#| eval: true
 #| warning: false
 #| message: false
+#| tbl-pos: H
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/speed_model.rds"))
+modelsummary(mean_speed_model)
 ```
 
+The coefficient estimates reveal that all predictors significantly contribute to the estimation of mean speed (`spd_mean`). The results indicate:
+
+Traffic Volume (`volume`): Higher traffic volumes are associated with slightly lower mean speeds, consistent with the expectation that congestion reduces average traffic speeds.
+5th Percentile Speed (`pct_05`): The lowest observed speeds strongly correlate with mean speed, suggesting a substantial influence of slower-moving vehicles on the overall mean.
+50th Percentile Speed (`pct_50`): The median speed demonstrates the largest positive relationship with mean speed, aligning with its central role in determining the distribution of traffic speeds.
+95th Percentile Speed (`pct_95`): Faster-moving vehicles also positively impact the mean speed, but the effect is less pronounced than that of the median speed.
+These results highlight the combined importance of different speed thresholds and traffic volume in shaping overall traffic behavior.
+
+## Actual vs. Predicted Mean Speeds {#sec-actualvspredicted}
+To evaluate the predictive performance of our Bayesian linear regression model, we compare the predicted mean speeds, derived from the model, with the actual observed mean speeds in the dataset.
+The comparison is visualized in @fig-actualvspredicted.
+
 ```{r}
+#| label: fig-actualvspredicted
+#| fig-cap: Comparison of actual and predicted mean speeds.
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
+#| fig-pos: H
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+filtered_data <- filtered_data %>%
+  mutate(
+    predicted = predict(mean_speed_model, newdata = filtered_data)
   )
+
+ggplot(filtered_data, aes(x = predicted, y = spd_mean)) +
+  geom_point(alpha = 0.5, color = "blue") +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(
+    x = "Predicted Mean Speeds",
+    y = "Actual Mean Speeds"
+  ) +
+  theme_minimal()
+
 ```
 
-# Discussion
+In @fig-actualvspredicted, each point represents an observation where the x-axis denotes the predicted mean speed, and the y-axis shows the actual mean speed.
+The red dashed line represents the ideal case where predictions perfectly match the actual values (i.e., a 45-degree reference line).
+Observations closely aligning with the dashed line indicate a strong agreement between predicted and actual values.
+
+The model demonstrates strong predictive accuracy, as the majority of points cluster tightly along the dashed line with minimal deviation.
+This close alignment suggests that the model captures the relationship between the predictors (`pct_05`, `pct_50`, `pct_95`, and `volume`) and the response variable (`spd_mean`) effectively, with minimal systematic bias.
+
+Notably, there are only a few points where slight deviations from the line are visible, mostly at higher mean speed values.
+This could be attributed to minor outliers or unique conditions not fully explained by the model's predictors.
+Overall, the results confirm the reliability of the Bayesian model in predicting mean traffic speeds, especially for the typical speed ranges observed in the dataset.
+
+We explore the broader implications of these results, including potential improvements or limitations in the model, in @sec-discussion.
+
+# Discussion {#sec-discussion}
 
@@
 
-pp_check(first_model) +
+pp_check(mean_speed_model) +
   theme_classic() +
@@
 
-posterior_vs_prior(first_model) +
+posterior_vs_prior(mean_speed_model) +
   theme_minimal() +
@@
 
-plot(first_model, "trace")
+plot(mean_speed_model, "trace")
 
-plot(first_model, "rhat")
+plot(mean_speed_model, "rhat") + 
+  theme_classic()
 ```

### 2024-12-03 paper/paper.qmd
update paper and datasheet

@@
   - Andy Jiang
-thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/toronto-vehicles-speed-analysis)."
+thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-vehicles-speed-analysis."
 date: today
@@
 
+The remainder of this paper is structured as follows.
+@sec-data discusses the raw data, cleaning process, and key variables of interest, offering visual and tabular representations of the dataset.
+@sec-model introduces and justifies the Bayesian linear regression model used to analyze the relationship between traffic volume, speed percentiles, and mean vehicle speed.
+@sec-results presents the findings of the model, highlighting significant trends and correlations within the data.
+Finally, @sec-discussion explores the broader implications of these results, offering insights into traffic behavior and potential avenues for future research.
+Additional information regarding data observation methodology is provided in [Appendix -@sec-dataobservation], and further model diagnostics are detailed in [Appendix -@sec-model-details].
+
 # Data {#sec-data}
 
+The data used in this paper came from the Open Data Toronto portal through the library opendatatoronto [@citeODT], specifically the Safety Zone Watch Your Speed Program – Monthly Summary [@SpeedDataCite].
 All data analysis was conducted using R [@citeR] with the support of the following packages: tidyverse [@tidyverse], here [@here], ggplot2 [@ggplot2], rstanarm [@rstanarm], kableExtra [@kableExtra], modelsummary [@modelsummary], dplyr [@dplyr], and lubridate [@lubridate].
@@
 - Speed bins (e.g., 0-4 km/h, 5-9 km/h) for detailed breakdowns of vehicle counts across speed ranges.
-- @tbl-speed-preview showcases a preview of the cleaned dataset with selected columns to highlight key features.
+- @tbl-speed-preview showcases a simplified preview of the dataset with selected columns to highlight key features.
 
@@
 #| label: tbl-speed-preview
-#| tbl-cap: Preview of cleaned traffic speed dataset.
+#| tbl-cap: Preview of analysis traffic speed dataset.
 #| echo: false
@@
 data |> 
-  select(month, pct_05, pct_50, pct_95, spd_30, spd_50, spd_100_and_above, volume) |> 
+  select(month, pct_05, pct_95, spd_100_and_above, volume) |> 
   head(10) |> 
   kable(
-    col.names = c("Month", "5th Percentile", "Median (50th Percentile)", "95th Percentile", 
-                  "Vehicles (30-34 km/h)", "Vehicles (50-54 km/h)", "Vehicles (>100 km/h)", "Total Volume"),
+    col.names = c("Month", "5th Percentile", "95th Percentile", "Vehicles (>100 km/h)", "Total Volume"),
     booktabs = TRUE
@@
   sample_n(10) |> 
-  select(month, pct_05, pct_50, pct_95, volume, spd_100_and_above) |> 
+  select(month, pct_05, pct_95, spd_100_and_above, volume) |> 
   kable(
-    col.names = c("Month", "5th Percentile", "Median (50th Percentile)", "95th Percentile", "Total Volume", "Vehicles (>100 km/h)"),
+    col.names = c("Month", "5th Percentile", "95th Percentile", "Vehicles (>100 km/h)", "Total Volume"),
     booktabs = TRUE
@@
   labs(
-    x = "Year",
+    x = "Time",
     y = "Speed (km/h)",
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-The Bayesian multiple linear regression model used in this paper predicts the mean speed (spd_mean) as a function of total traffic volume (volume) and speed percentiles (pct_05, pct_50, and pct_95).
+The Bayesian multiple linear regression model used in this paper predicts the mean speed (`spd_mean`) as a function of total traffic volume (`volume`) and speed percentiles (`pct_05`, `pct_50`, and `pct_95`).
 The model is defined as follows:
@@
 
-- $\mu_i$ is the predicted mean speed (spd_mean) for the $i$-th observation given the total traffic volume and speed percentiles.
+- $\mu_i$ is the predicted mean speed (`spd_mean`) for the $i$-th observation given the total traffic volume and speed percentiles.
 - $\beta_0$ is the coefficient for the intercept.
-- $\beta_1$ is the coefficient for the predicted change in the mean speed given a one-unit increase in traffic volume (volume).
-- $\beta_2$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 5th percentile speed (pct_05).
-- $\beta_3$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 50th percentile speed (pct_50).
-- $\beta_4$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 95th percentile speed (pct_95).
+- $\beta_1$ is the coefficient for the predicted change in the mean speed given a one-unit increase in traffic volume (`volume`).
+- $\beta_2$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 5th percentile speed (`pct_05`).
+- $\beta_3$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 50th percentile speed (`pct_50`).
+- $\beta_4$ is the coefficient for the predicted change in the mean speed given a one-unit increase in the 95th percentile speed (`pct_95`).
 - $\sigma$ is the standard deviation of the residuals, representing unexplained variability in the mean speed.
@@
 
-# Results
+# Results {#sec-results}
 
-## Model Coefficients {#sec-modelcoefficients}
-The results of the Bayesian linear regression model estimating the relationship between mean traffic speed (spd_mean) and its key predictors are summarized in @tbl-modelsummary.
+## Model Coefficients
+The Bayesian linear regression model results, summarized in @tbl-modelsummary, provide insight into the relationship between mean traffic speed (`spd_mean`) and key predictors such as speed percentiles and traffic volume.
 
@@
 #| tbl-cap: Model summary of the predicted impact of traffic volume and speed percentiles on mean speed.
+#| eval: false
 #| echo: false
@@
 
-The coefficient estimates reveal that all predictors significantly contribute to the estimation of mean speed (`spd_mean`). The results indicate:
+The coefficient estimates reveal the following relationships:
 
-Traffic Volume (`volume`): Higher traffic volumes are associated with slightly lower mean speeds, consistent with the expectation that congestion reduces average traffic speeds.
-5th Percentile Speed (`pct_05`): The lowest observed speeds strongly correlate with mean speed, suggesting a substantial influence of slower-moving vehicles on the overall mean.
-50th Percentile Speed (`pct_50`): The median speed demonstrates the largest positive relationship with mean speed, aligning with its central role in determining the distribution of traffic speeds.
-95th Percentile Speed (`pct_95`): Faster-moving vehicles also positively impact the mean speed, but the effect is less pronounced than that of the median speed.
-These results highlight the combined importance of different speed thresholds and traffic volume in shaping overall traffic behavior.
+- Intercept (`-0.881`): This represents the baseline adjustment to the mean speed when all predictors are at their reference levels.
+- Traffic Volume (`0.000`): The coefficient indicates no measurable impact of total traffic volume on mean speed, which is unexpected given traditional congestion dynamics.
+- 5th Percentile Speed (`0.206`): Lower speeds positively influence the mean, showing that slower-moving vehicles contribute moderately to the overall average.
+- 50th Percentile Speed (`0.652`): Median speeds have the largest positive effect on mean speed, emphasizing their central role in shaping the distribution of traffic speeds.
+- 95th Percentile Speed (`0.184`): Higher speeds have a smaller but still positive influence on mean speed, reflecting their contribution to the upper end of the distribution.
 
-## Actual vs. Predicted Mean Speeds {#sec-actualvspredicted}
+The model's performance metrics underscore its robustness:
+
+- R² and Adjusted R² (`0.992`): The model explains 99.2% of the variance in mean speed, indicating excellent fit.
+- Log-Likelihood (`-41414.100`): Represents the model's likelihood given the data.
+- WAIC and LOOIC (`82872.9`): These information criteria indicate model performance, balancing goodness-of-fit and model complexity.
+- RMSE (`0.88`): A low root mean square error confirms the model's predictive accuracy.
+
+These results highlight the dominant role of speed percentiles in determining mean speed while raising questions about the negligible impact of traffic volume.
+Further exploration into potential interactions or omitted variables could clarify this finding.
+
+## Actual vs. Predicted Mean Speeds
 To evaluate the predictive performance of our Bayesian linear regression model, we compare the predicted mean speeds, derived from the model, with the actual observed mean speeds in the dataset.
@@
 
-## First discussion point {#sec-first-point}
+## Post-pandemic trends and implications for traffic management {#sec-postpandemic}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The analysis in @fig-extreme-speeds-vs-volume illustrates the significant divergence in traffic patterns between the pre- and post-2020 periods.
+Post-pandemic traffic volumes are notably higher and are accompanied by an increase in extreme speeds (vehicles traveling over 100 km/h).
+This trend may reflect the rapid population growth in Toronto, driven by immigration, alongside shifts in urban mobility post-COVID-19.
 
-## Second discussion point
+The steeper slope for post-2020 data in @fig-extreme-speeds-vs-volume suggests that extreme speed counts have become more sensitive to increases in traffic volume.
+One potential explanation for this is the stress placed on infrastructure as more drivers navigate an increasingly dense urban network.
+These findings highlight the importance of re-evaluating existing traffic management strategies and prioritizing policies that address the dangers posed by extreme speeds in high-traffic environments.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Targeted interventions, such as extending safety zones, introducing stricter enforcement of speed limits, or implementing dynamic speed regulation systems, could help mitigate the risks associated with this upward trend in extreme speeds.
+Additionally, these findings underscore the importance of accounting for macro-level disruptions, like the pandemic, in transportation policy design to better anticipate and address evolving urban mobility challenges.
 
-## Third discussion point
+## Variability in speed percentiles and safety implications {#sec-variability}
 
-## Weaknesses and next steps
+The trends in speed percentiles visualized in @fig-key-percentiles reveal nuanced changes in traffic behavior over time.
+While the 95th percentile speed has shown a steady decline to below 50 km/h, suggesting progress in limiting higher speeds, the sharp drop in the 5th percentile speed since mid-2022 warrants attention.
+A decline from approximately 10 km/h to 5 km/h may indicate increased congestion, construction zones, or other road conditions causing slower-moving traffic.
+This could create a wider speed differential among vehicles, potentially leading to more frequent and severe accidents.
 
-Weaknesses and next steps should also be included.
+The stability of the 50th percentile speed at approximately 30 km/h reflects consistency in typical driving conditions, while the steady spread between the 5th and 95th percentiles (35–40 km/h) points to minimal change in speed variability.
+However, the importance of addressing the outliers at the extremes cannot be overstated, as they have disproportionate safety implications.
+This analysis supports the need for targeted measures to minimize variability in speeds, including enhanced road design and enforcement strategies to reduce abrupt accelerations or decelerations.
+
+## Limitations in modeling and data interpretation {#sec-limitations}
+
+While the Bayesian linear regression model employed in this paper provides meaningful insights into the relationship between key predictors and mean vehicle speed, several limitations warrant discussion.
+As highlighted in @sec-model, the model does not account for potential interactions among predictors, such as the interplay between traffic volume and percentile speeds.
+For instance, higher traffic volumes may disproportionately affect the 5th percentile speed compared to the 95th percentile, but this relationship remains unexplored due to the model’s linear structure.
+
+Additionally, as noted in @sec-data, the dataset’s reliance on radar-equipped safety zone signs introduces biases related to location and driver behavior.
+Vehicles may reduce speed upon approaching these signs, potentially underestimating typical speed distributions for the broader road network.
+Moreover, the dataset aggregates observations monthly, which may obscure short-term fluctuations in traffic patterns.
+
+Future modeling efforts could incorporate interaction terms and hierarchical structures to better capture the complexity of traffic dynamics.
+Similarly, integrating data from diverse monitoring systems (e.g., highway cameras or GPS datasets) could provide a more comprehensive view of Toronto’s traffic behavior across different contexts.
+
+## Policy recommendations and future directions {#sec-policy}
+
+The findings in this paper provide a valuable foundation for improving traffic safety and efficiency in Toronto. Several actionable recommendations emerge:
+
+- Enhance enforcement and safety measures for extreme speeds: As post-2020 trends demonstrate a growing prevalence of extreme speeds, prioritizing enforcement in high-traffic areas could reduce associated risks.
+The use of automated speed cameras, stricter penalties for excessive speeding, and expanded safety zone coverage are potential solutions.
+
+- Address speed variability through road design: Consistent median speeds and limited speed spreads are desirable for traffic flow and safety.
+Urban planning efforts should focus on designing roads that naturally discourage extreme speeds while avoiding conditions that create bottlenecks for slower vehicles.
+
+- Incorporate adaptive traffic policies: The sensitivity of extreme speeds to traffic volume post-2020 underscores the importance of adaptive measures.
+Dynamic speed limits that adjust based on real-time traffic and weather conditions could help balance safety and mobility.
+
+- Explore complementary data sources: Expanding beyond radar-collected data to include GPS-based datasets, time-of-day analyses, and seasonal effects could provide richer insights.
+Doing so would allow for a deeper understanding of the factors driving variability in traffic behavior.
+
+These recommendations highlight the broader applicability of the study’s findings to urban centers experiencing rapid population growth and evolving mobility challenges.
+By addressing the identified gaps and leveraging the insights gained, Toronto and similar cities can move toward safer and more efficient transportation systems.
 
@@
 
-# Additional data details
+# Data Observation {#sec-dataobservation}
+
+## Idealized Methodology for Traffic Data Collection
+The observational data analyzed in this study stems from the City of Toronto’s Safety Zone Watch Your Speed Program (WYSP), where radar-equipped speed signs monitor and aggregate vehicle speeds.
+While this data provides valuable insights, it also introduces limitations such as restricted spatial and temporal coverage, potential behavioral biases due to the visibility of speed signs, and an inability to capture all traffic volume accurately.
+This section proposes an idealized methodology to address these limitations and enhance the robustness of traffic speed data collection.
+
+### Goals of the Idealized Methodology
+
+The idealized methodology aims to:
+
+- Expand Representativeness: Ensure the data reflects traffic patterns across different locations, road types, and times of the day.
+- Minimize Behavioral Biases: Reduce the influence of drivers’ behavior changes caused by the visibility of monitoring devices.
+- Capture Temporal Variability: Account for daily, seasonal, and situational fluctuations in traffic patterns.
+- Improve Accuracy: Ensure that all traffic, including vehicles that might evade radar detection, is measured consistently.
+
+### Proposed Framework
+- Stratified Sampling Approach
+  - To improve spatial representativeness, Toronto can be divided into strata based on geographic regions (e.g., downtown, suburban, and rural areas) and road types (e.g., highways, arterial roads, local streets).
+The number of monitoring devices deployed in each stratum should be proportional to traffic density and road importance.
+This stratified approach ensures adequate representation of all areas and road conditions.
+
+- Rotational Monitoring
+  - Rather than using permanent installations, monitoring devices should be deployed on a rotational basis.
+Devices can be moved periodically across selected sites within each stratum, ensuring broader spatial coverage.
+Rotation would also reduce drivers' habituation to the presence of speed signs, which might otherwise alter their driving behavior.
+
+- Integration of Covert Monitoring
+  - To minimize behavioral biases, the methodology should integrate covert monitoring techniques.
+Hidden radar devices or automated cameras can measure vehicle speeds without drivers being aware of their presence, thus capturing more natural driving behavior.
+
+- Temporal Sampling
+  - Data collection should account for temporal variability by monitoring during different times of day (e.g., peak vs.
+off-peak hours), days of the week (e.g., weekdays vs.
+weekends), and seasons.
+This approach ensures a comprehensive understanding of traffic patterns and their fluctuations.
+
+- Technological Enhancements
+  - Advanced technologies such as LiDAR and AI-based sensors can supplement traditional radar devices to increase measurement accuracy.
+These technologies can detect multiple vehicles simultaneously, reducing blind spots and capturing speeds across a wider field of view.
+
+- Calibration and Standardization
+  - All monitoring devices must undergo regular calibration to maintain consistency and accuracy in speed measurements.
+A centralized standardization protocol should be implemented to ensure uniformity across all collected data.
+
+### Challenges and Practical Considerations
+Implementing this idealized methodology involves logistical and financial challenges:
+
+- Cost Implications: Deploying advanced devices and rotational monitoring systems requires significant investment in equipment and personnel.
+- Coordination with Stakeholders: Accessing certain road segments for device installation may require coordination with local authorities, businesses, and residents.
+- Data Management: Large-scale data collection introduces challenges related to storage, integration, and processing. Ensuring data integrity and consistency is critical for effective analysis.
+
+Despite these challenges, the proposed framework offers a robust and realistic solution for improving traffic data collection.
+By addressing the limitations of the existing WYSP dataset, this methodology provides a foundation for more accurate and actionable insights into urban mobility.
+
+## Simulation for Evaluating the Proposed Framework
+Simulations can be an effective way to evaluate the proposed methodology before large-scale implementation. A simulation study could:
+
+- Model Traffic Flow: Use agent-based modeling to simulate traffic flow across Toronto, incorporating variables such as road type, time of day, and vehicle characteristics.
+- Test Monitoring Scenarios: Evaluate the effectiveness of rotational and covert monitoring setups in capturing accurate speed data across different strata.
+Assess Behavioral Impact: Simulate changes in driver behavior due to visible and hidden monitoring devices to quantify potential biases.
+- Optimize Resource Allocation: Simulate various deployment strategies to determine the optimal allocation of monitoring devices for maximum coverage and cost efficiency.
+
+By simulating these scenarios, stakeholders can refine the methodology and address potential challenges before field deployment.
+
+## Strengths and Limitations of Observational Data
+The WYSP dataset offers valuable observational insights into Toronto’s traffic patterns, but it is not without its limitations.
+
+### Strengths
+- Ease of Collection: Radar speed signs provide a non-intrusive method for collecting large amounts of data over extended periods.
+- Focus on Safety Zones: Data collection in designated Safety Zones highlights areas with specific traffic and safety concerns.
+- Detailed Metrics: The dataset includes percentile-based speed metrics (e.g., 5th, 50th, and 95th percentiles), which are critical for analyzing traffic behavior.
+
+### Limitations
+- Limited Spatial Coverage: The dataset only includes areas with installed speed signs, omitting many road segments across Toronto.
+- Behavioral Biases: Visible speed signs may influence drivers to temporarily reduce their speed, resulting in data that may not reflect typical behavior.
+- Temporal Gaps: Data is aggregated monthly, obscuring finer temporal variations such as hourly or daily patterns.
+- Non-Representative Traffic Volume: The count of vehicles measured by speed signs is not equivalent to total traffic volume, as some vehicles may evade detection.
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+The posterior predictive check (PPC) evaluates how well the Bayesian model fits the observed data [@tellingstories].
+In @fig-ppcheck, the black line represents the observed data density, while the blue lines represent the density of replicated data generated from the posterior predictive distribution.
+The alignment between these lines indicates the model's ability to reproduce the data.
+While there is a slight deviation at the peak of the density curve, the overall alignment suggests that the model provides a reasonable fit to the observed data.
+Minor discrepancies may warrant adjustments in the model specification or inclusion of additional predictors to improve fit.
 
 ```{r}
-#| eval: true
+#| label: fig-ppcheck
+#| fig-cap: "Examining how the model fits the data through a posterior predictive check"
 #| echo: false
+#| fig-pos: H
 #| message: false
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
@@
   theme(legend.position = "bottom")
+
+```
+
+## Comparison of the Posterior and Prior
+@fig-postprior compares the prior and posterior distributions for the model parameters.
+The posterior (left) is informed by the observed data, while the prior (right) reflects initial assumptions [@tellingstories].
+Notably, the posterior for all parameters, except the intercept and sigma, remains centered around zero, showing minimal change from the prior.
+The sigma parameter has shifted slightly to the right, reflecting a more informed estimate based on the data.
+The red intercept dot spans the entire horizontal range in the prior, indicating a vague prior assumption.
+This comparison illustrates that the data provides substantial information for refining parameter estimates.
+
+```{r}
+#| label: fig-postprior
+#| fig-cap: "Comparison of posterior and prior distributions for model parameters"
+#| echo: false
+#| warning: false
+#| fig-pos: H
+#| message: false
 
@@
   coord_flip()
+
 ```
 
-## Diagnostics
+## Markov Chain Monte Carlo Convergence
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
+### Trace Plot
+The trace plot in @fig-trace examines the behavior of the MCMC chains for each parameter [@tellingstories].
+The chains exhibit stable horizontal fluctuations with no discernible trends, indicating proper mixing and convergence.
+The patterns across all six parameters are consistent, suggesting that the algorithm has adequately explored the parameter space.
 ```{r}
+#| label: fig-trace
+#| fig-cap: "Trace plot for MCMC convergence assessment"
 #| echo: false
-#| eval: true
 #| message: false
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| fig-pos: H
 
@@
 
+```
+
+### Rhat Plot
+The Rhat plot in @fig-rhat evaluates the convergence of the Markov Chain Monte Carlo (MCMC) algorithm.
+The Rhat values for all parameters hover just above 1.00, indicating acceptable convergence.
+While `pct_05` displays the longest interval, it still falls within the acceptable range.
+This diagnostic suggests that the chains for all parameters have reached convergence and that the posterior samples are reliable.
+
+```{r}
+#| label: fig-rhat
+#| fig-cap: "Rhat plot for MCMC convergence assessment"
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-pos: H
+
 plot(mean_speed_model, "rhat") +
+  theme_classic()
+
+```
+
+## Credible Intervals
+The credible intervals for the model coefficients are shown in @fig-credible-intervals.
+The intercept is negative, suggesting a downward adjustment for baseline mean speed.
+The `volume` coefficient is centered at 0, indicating little influence on the predicted mean speed.
+Among the predictors, `pct_50` has the strongest positive effect, with a credible interval between 0.5 and 1.0, while `pct_05` and `pct_95` show moderate positive effects.
+The residual standard deviation (`sigma`) is near 1.0, reflecting the model's uncertainty in capturing variability.
+```{r}
+#| label: fig-credible-intervals
+#| fig-cap: "95% Credible Intervals for Model Coefficients"
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-pos: H
+
+plot(mean_speed_model) +
   theme_classic()

### 2024-12-03 other/datasheet/datasheet.qmd
update datasheet path

---
title: "Datasheet for 'Toronto Traffic Speed Analysis Dataset'"
author: 
  - Andy Jiang
thanks: "Code and data are available at: https://github.com/AndyYanxunJiang/toronto-vehicles-speed-analysis."
date: today
date-format: long
abstract: "Provides access to traffic speed data collected in Toronto's designated safety zones between 2017 and 2024. This dataset enables analysis of urban mobility, safety trends, and the relationship between traffic volume and speed percentiles. This datasheet enhances the reproducibility and understanding of research conducted using this dataset."
format: pdf
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to monitor traffic speeds, enforce road safety, and provide insights into traffic behavior in designated Safety Zones in Toronto. It allows for analyzing trends and assessing traffic management strategies.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by the City of Toronto as part of the Safety Zone Watch Your Speed Program (WYSP) [@SpeedDataCite].
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The dataset is funded by the City of Toronto as part of its traffic safety initiatives.
4. *Any other comments?*
    - N/A

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each row represents aggregated monthly traffic data collected from radar-equipped Safety Zone speed signs, including speed percentiles, vehicle counts in speed bins, and total traffic volumes.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset contains 31,990 observations (monthly data from 2017–2024).
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- It is a sample limited to traffic monitored in specific Safety Zones across Toronto.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance includes aggregated speed and volume data, such as 5th, 50th, and 95th percentile speeds, and counts of vehicles in predefined speed ranges.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- No labels or targets, but each instance is timestamped (monthly).
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some speed bins may have zero counts, reflecting low or no traffic in specific ranges during certain months.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No explicit relationships; each observation is independent and time-indexed.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Potential behavioral biases exist due to the visibility of speed signs, which may cause drivers to slow down.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.
16. *Any other comments?*
	- N/A

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data was collected using radar-equipped speed signs installed in Toronto Safety Zones [@SpeedDataCite].
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Radar speed signs measure oncoming vehicle speeds with an accuracy of ±1 km/h and aggregate the results monthly [@SpeedDataCite].
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Locations for monitoring were selected by the City of Toronto, focusing on Safety Zones.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- City of Toronto employees.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- Data was collected monthly from 2017–2024.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Data was collected automatically through radar speed signs.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Drivers passing by signs may be aware of data collection based on posted signage.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Not required; no personal information was collected.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
12. *Any other comments?*
	- N/A

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Data was aggregated and cleaned to remove erroneous or incomplete observations.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- N/A
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- N/A
4. *Any other comments?*
	- N/A

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, for analyzing trends in traffic speed and volume, as detailed in this paper.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other acess point.*
	- No known repository beyond the Open Data Toronto portal.
3. *What (other) tasks could the dataset be used for?*
	- Studying seasonal variations in traffic, evaluating traffic policy impacts, or identifying high-risk zones for accidents.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- The dataset’s focus on Safety Zones may limit its generalizability to other urban or rural areas.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used to generalize beyond the specific Safety Zones it covers.
6. *Any other comments?*
	- N/A

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, through the Open Data Toronto portal.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- As a CSV file or through an API provided by Open Data Toronto.
3. *When will the dataset be distributed?*
	- It is available now.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Yes, as per the Open Data Toronto terms of use.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.
7. *Any other comments?*
	- N/A

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Open Data Toronto.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Through the Open Data Toronto portal.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- Yes, as new data is collected by the City of Toronto.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- N/A
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Unknown, but likely updated periodically.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- The dataset is available through the Open Data Toronto portal. Users who wish to augment or contribute to the dataset must contact the City of Toronto's Open Data team via their official website (https://open.toronto.ca). Contributions are typically validated by the data custodians to ensure consistency with municipal data standards.
8. *Any other comments?*
	- N/A

\newpage
# Reference

### 2024-12-03 paper/paper.qmd
update README and clean paper
update reference and evaluate modelsummary

@@
 ---
-title: "Analyzing Toronto Traffic Speeds: Insights into Patterns and Predictive Modeling"
-subtitle: "Bayesian Linear Regression Reveals Key Determinants of Mean Speed and Post-Pandemic Trends"
+title: "Analyzing Toronto Traffic Speeds: Insights into Urban Mobility, Safety, and Speed Dynamics"
+subtitle: "Exploring Key Traffic Patterns and Relationships Between Speed Percentiles and Volume"
 author: 
@@
 # Introduction
+
 Traffic speed analysis is a critical tool for understanding urban mobility, road safety, and congestion.
@@
 ## Overview
+
 This paper analyzes traffic speed data collected in Toronto over the period 2017–2024.
@@
 ## Speed Dataset Preview
+
 The traffic speed dataset contains information collected from multiple locations and months, detailing percentile speeds, total traffic volume, and counts of vehicles traveling within specific speed ranges.
@@
 This dataset includes:
+
 - Percentile speeds (5th, 10th, ... 95th) representing speed thresholds exceeded by corresponding proportions of vehicles.
@@
 ## Random Sample from the Dataset
+
 @tbl-random-sample provides a random sample of 10 observations, offering an unbiased snapshot of key metrics such as speed percentiles, total traffic volume, and counts of vehicles exceeding 100 km/h.
@@
 ## Extreme Speeds Dataset
+
 Vehicles traveling at extreme speeds (over 100 km/h) represent a critical factor in understanding traffic safety and violations.
@@
 ## Proportion of Vehicles at Moderate Speeds
+
 Another key insight from this dataset is the proportion of vehicles traveling within a moderate speed range (50–70 km/h).
 This measure helps understand how typical driving speeds align with safe and efficient traffic flow.
-
 @tbl-speed-proportion highlights monthly proportions of vehicles in this speed range.
@@
 
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
 # Model {#sec-model}
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of [@rstanarm]. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package [@rstanarm].
 
@@
 ## Model Coefficients
+
 The Bayesian linear regression model results, summarized in @tbl-modelsummary, provide insight into the relationship between mean traffic speed (`spd_mean`) and key predictors such as speed percentiles and traffic volume.
@@
 #| tbl-cap: Model summary of the predicted impact of traffic volume and speed percentiles on mean speed.
-#| eval: false
 #| echo: false
@@
 ## Actual vs. Predicted Mean Speeds
+
 To evaluate the predictive performance of our Bayesian linear regression model, we compare the predicted mean speeds, derived from the model, with the actual observed mean speeds in the dataset.
@@
 
-## Post-pandemic trends and implications for traffic management {#sec-postpandemic}
+## Post-pandemic trends and implications for traffic management
 
@@
 
-## Variability in speed percentiles and safety implications {#sec-variability}
+## Variability in speed percentiles and safety implications
 
@@
 
-## Limitations in modeling and data interpretation {#sec-limitations}
+## Limitations in modeling and data interpretation
 
@@
 
-## Policy recommendations and future directions {#sec-policy}
+## Challenges and Opportunities in Traffic Enforcement
 
-The findings in this paper provide a valuable foundation for improving traffic safety and efficiency in Toronto. Several actionable recommendations emerge:
+Automated enforcement systems present both significant opportunities and challenges for traffic management in Toronto.
+Evidence from cities like Washington, D.C., and New York City highlights the potential of these systems to address traffic violations effectively.
+For example, Washington, D.C.'s bus-mounted cameras issued over 147,000 tickets to drivers illegally using bus-only lanes within months of implementation, showcasing their efficiency in catching violations that might otherwise go unnoticed.
+Similarly, New York's Automated Bus Lane Enforcement (ABLE) program has been expanded to address issues like double parking, further demonstrating the adaptability of automated systems [@Elliott_2024].
 
-- Enhance enforcement and safety measures for extreme speeds: As post-2020 trends demonstrate a growing prevalence of extreme speeds, prioritizing enforcement in high-traffic areas could reduce associated risks.
-The use of automated speed cameras, stricter penalties for excessive speeding, and expanded safety zone coverage are potential solutions.
+In Toronto, while automated speed enforcement (ASE) cameras have successfully reduced speeding, the city's adoption of similar technology for other traffic violations, such as blocking intersections ("blocking the box") or driving in transit lanes, has been limited.
+This is partly due to legislative constraints at the provincial level, which prevent issuing fines for certain violations if the vehicle owner was not the driver at the time.
+Additionally, despite the proven efficacy of ASE cameras in reducing vehicle speeds, as noted by the significant decrease in speeding incidents in school zones, the rollout of new cameras has been slow.
 
-- Address speed variability through road design: Consistent median speeds and limited speed spreads are desirable for traffic flow and safety.
-Urban planning efforts should focus on designing roads that naturally discourage extreme speeds while avoiding conditions that create bottlenecks for slower vehicles.
+These delays underscore broader challenges in traffic enforcement, including bureaucratic hurdles, resource constraints, and reliance on traditional enforcement methods.
+Moreover, the cultural perspective within traffic policing, which often deprioritizes ticketing for non-criminal violations, exacerbates these issues.
+However, these challenges also present opportunities for leveraging technology to complement traditional enforcement and create a more comprehensive traffic management framework.
 
-- Incorporate adaptive traffic policies: The sensitivity of extreme speeds to traffic volume post-2020 underscores the importance of adaptive measures.
-Dynamic speed limits that adjust based on real-time traffic and weather conditions could help balance safety and mobility.
+## Policy recommendations and future directions
 
-- Explore complementary data sources: Expanding beyond radar-collected data to include GPS-based datasets, time-of-day analyses, and seasonal effects could provide richer insights.
-Doing so would allow for a deeper understanding of the factors driving variability in traffic behavior.
+The findings in this study offer a foundation for improving traffic safety and efficiency in Toronto.
+A central takeaway is the value of automated enforcement in addressing critical challenges, including extreme speeds, variability in traffic patterns, and enforcement gaps.
+Automated speed enforcement (ASE) cameras have already demonstrated significant success in Toronto, reducing both the prevalence of speeding and overall vehicle speeds in monitored areas.
+Expanding the use of these cameras and integrating new functionalities, such as detecting transit lane violations or intersection blockages, can further enhance their impact.
+Learning from other cities like Washington, D.C., and New York, Toronto can adopt similar innovations to target a broader range of traffic offenses [@Elliott_2024].
 
-These recommendations highlight the broader applicability of the study’s findings to urban centers experiencing rapid population growth and evolving mobility challenges.
-By addressing the identified gaps and leveraging the insights gained, Toronto and similar cities can move toward safer and more efficient transportation systems.
+Urban road design remains another critical area for intervention.
+Consistent median speeds and limited speed variability are essential for maintaining safe and efficient traffic flow.
+To achieve this, urban planning efforts should focus on designing roads that naturally discourage dangerous speeds, while avoiding features that create bottlenecks or increase risks for vulnerable road users.
+For instance, implementing physical design elements such as raised crosswalks or lane-narrowing strategies can complement automated enforcement measures.
+
+Adaptive traffic policies also hold promise for addressing the dynamic nature of urban mobility.
+Introducing technologies that adjust speed limits in real time based on traffic conditions, weather, or specific safety concerns could help balance safety with mobility.
+By integrating ASE devices into such adaptive frameworks, the city can enhance their effectiveness in high-risk or high-traffic areas.
+
+To support these efforts, data collection and analysis must be expanded.
+Toronto can benefit from leveraging complementary data sources, such as GPS datasets, time-of-day analyses, and seasonal traffic patterns.
+Such data can refine the deployment of automated enforcement and inform broader policy decisions.
+For example, cities like Washington, D.C., have used bus-mounted cameras to monitor transit lane violations effectively.
+Toronto could similarly implement pilot programs to gather data and address challenges such as "blocking the box" and illegal lane usage.
+
+Finally, addressing legislative barriers is essential for unlocking the full potential of automated enforcement.
+While current laws limit certain applications, proactive steps—such as deploying cameras to collect data even before legislative approval for expanded ticketing—can help build the case for broader implementation.
+The city's leaders should take bold steps, pushing forward with camera deployment to enhance safety and traffic management, even if provincial approvals are delayed.
+This approach, modeled after prior successes in advocating for speed cameras, could expedite the necessary legislative changes.
+
+By integrating automated enforcement, adaptive policies, and data-driven design into a cohesive strategy, Toronto can tackle its pressing traffic challenges and create safer, more efficient urban mobility systems.
 
@@
 ## Idealized Methodology for Traffic Data Collection
+
 The observational data analyzed in this study stems from the City of Toronto’s Safety Zone Watch Your Speed Program (WYSP), where radar-equipped speed signs monitor and aggregate vehicle speeds.
@@
 ### Proposed Framework
+
 - Stratified Sampling Approach
@@
 ### Challenges and Practical Considerations
+
 Implementing this idealized methodology involves logistical and financial challenges:
@@
 ## Simulation for Evaluating the Proposed Framework
+
 Simulations can be an effective way to evaluate the proposed methodology before large-scale implementation. A simulation study could:
@@
 
+## Automated Speed Enforcement (ASE) Evaluation in Toronto
+
+An evaluation of three years of data from Toronto's Automated Speed Enforcement (ASE) cameras provides valuable insights into the potential of automated systems to influence driver behavior and enhance traffic safety.
+Conducted by researchers from SickKids Hospital and Toronto Metropolitan University, this study highlights the effectiveness of ASE in reducing speeding across multiple speed zones [@Rodrigues_2023].
+
+From January 2020 to December 2022, the City of Toronto installed 50 ASE cameras, rotating them to different locations periodically, and later added 25 more cameras to increase coverage.
+The study reported the following significant findings:
+
+- Reduction in Speeding: The proportion of vehicles speeding decreased by 45% in areas with an ASE device. In specific zones:
+  - 30 km/h zones: Speeding dropped from 60% to 43%.
+  - 40 km/h zones: Speeding dropped from 51% to 30%.
+  - 50 km/h zones: Speeding dropped from 58% to 36%.
+
+- Decrease in Operating Speeds: Vehicle operating speeds decreased by approximately 7 km/h overall in areas with ASE cameras:
+  - 30 km/h zones: From 44 to 37 km/h.
+  - 40 km/h zones: From 50 to 44 km/h.
+  - 50 km/h zones: From 63 to 60 km/h.
+
+- Reduction in Excessive Speeds: The proportion of drivers exceeding speed limits by 20 km/h or more dropped by 87% following ASE implementation.
+
+- Widespread Impact: Speed limit compliance improved at 80% of locations equipped with ASE devices.
+
+This data demonstrates ASE's substantial role in curbing excessive speeding and creating safer road environments, particularly near schools and residential areas.
+The findings align with broader efforts to design data-driven traffic safety policies and underscore the importance of integrating automated systems into traffic management strategies.
+
+### Implications for Observational Data Methodology
+
+The ASE program also reflects several elements of an idealized data collection methodology:
+
+- Rotational Deployment: Cameras were rotated across locations to broaden coverage and minimize behavioral biases from static monitoring.
+- Comprehensive Data Collection: Despite some limitations imposed by the COVID-19 pandemic, the study captured substantial data during periods of normal activity, enhancing its reliability.
+- Focused Interventions: Placement of ASE cameras in high-risk areas, such as school zones, demonstrated a targeted approach to improving road safety.
+
+This case study provides a model for integrating automated enforcement with observational data collection, highlighting its efficacy in reducing risky driving behaviors and promoting safer urban mobility.
+
 ## Strengths and Limitations of Observational Data
+
 The WYSP dataset offers valuable observational insights into Toronto’s traffic patterns, but it is not without its limitations.
@@
 ### Strengths
+
 - Ease of Collection: Radar speed signs provide a non-intrusive method for collecting large amounts of data over extended periods.
@@
 ### Limitations
+
 - Limited Spatial Coverage: The dataset only includes areas with installed speed signs, omitting many road segments across Toronto.
@@
 ## Comparison of the Posterior and Prior
+
 @fig-postprior compares the prior and posterior distributions for the model parameters.
@@
 ### Trace Plot
+
 The trace plot in @fig-trace examines the behavior of the MCMC chains for each parameter [@tellingstories].
@@
 The patterns across all six parameters are consistent, suggesting that the algorithm has adequately explored the parameter space.
+
 ```{r}
@@
 ### Rhat Plot
+
 The Rhat plot in @fig-rhat evaluates the convergence of the Markov Chain Monte Carlo (MCMC) algorithm.
@@
 ## Credible Intervals
+
 The credible intervals for the model coefficients are shown in @fig-credible-intervals.

### 2024-12-03 other/datasheet/datasheet.qmd
update datasheet

@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The dataset was created to monitor traffic speeds, enforce road safety, and provide insights into traffic behavior in designated Safety Zones in Toronto.
+    - The dataset was created to monitor traffic speeds, enforce road safety, and provide understandings of traffic behavior in designated Safety Zones in Toronto.
 It allows for analyzing trends and assessing traffic management strategies.

### 2024-12-03 paper/paper.qmd
finalize paper

@@
 ---
-title: "Analyzing Toronto Traffic Speeds: Insights into Urban Mobility, Safety, and Speed Dynamics"
-subtitle: "Exploring Key Traffic Patterns and Relationships Between Speed Percentiles and Volume"
+title: "Analyzing Toronto Traffic Speeds: Understanding Urban Mobility, Safety, and Speed Dynamics"
+subtitle: "Exploring Traffic Patterns and Relationships Between Speed Percentiles and Volume"
 author: 
@@
 Results highlight significant changes in traffic patterns post-2020, with increased extreme speeds linked to higher traffic volumes.
-This study provides insights into urban mobility dynamics, offering a basis for traffic management strategies and policy-making."
+This study provides a deeper understanding of urban mobility dynamics, supporting traffic management strategies and policy-making."
 format: pdf
@@
 
-Traffic speed analysis is a critical tool for understanding urban mobility, road safety, and congestion.
+Traffic speed analysis is an essential method for understanding urban mobility, road safety, and congestion.
 In metropolitan areas like Toronto, shifting traffic patterns reflect changes in urban infrastructure, population growth, and behavioral adaptations to external events, such as the COVID-19 pandemic.
@@
 The dataset comprises monthly summaries of speed percentiles, vehicle counts across specific speed ranges, and total traffic volumes.
-Notably, the study explores trends in extreme speeds (vehicles exceeding 100 km/h) and evaluates how they have evolved before and after 2020, a period marked by rapid population growth and post-pandemic recovery.
+Notably, the study examines trends in extreme speeds (vehicles exceeding 100 km/h) and evaluates how they have changed before and after 2020, a period marked by rapid population growth and post-pandemic recovery.
 These analyses aim to address a key gap: understanding how percentile-based speed measures and traffic volume influence overall mean speeds.
@@
 @sec-results presents the findings of the model, highlighting significant trends and correlations within the data.
-Finally, @sec-discussion explores the broader implications of these results, offering insights into traffic behavior and potential avenues for future research.
+Finally, @sec-discussion explores the broader implications of these results, offering understandings of traffic behavior and potential avenues for future research.
 Additional information regarding data observation methodology is provided in [Appendix -@sec-dataobservation], and further model diagnostics are detailed in [Appendix -@sec-model-details].
@@
 
-The dataset aggregates metrics monthly, offering insights into how traffic behavior has changed over time.
+The dataset aggregates metrics monthly, offering understandings of how traffic behavior has changed over time.
 This analysis is particularly focused on extreme speeds (vehicles exceeding 100 km/h) and their relationship with total traffic volume, as well as trends in key speed percentiles.
@@
 Furthermore, the data is limited to Safety Zones where the speed signs have been installed, restricting its spatial coverage.
-Despite these limitations, the dataset is well-documented, complete, and up-to-date, providing valuable insights into traffic speed patterns across Toronto.
+Despite these limitations, the dataset is well-documented, complete, and up-to-date, providing useful understandings of traffic speed patterns across Toronto.
 
@@
 The traffic speed dataset contains information collected from multiple locations and months, detailing percentile speeds, total traffic volume, and counts of vehicles traveling within specific speed ranges.
-The dataset aggregates these metrics monthly, offering insights into traffic behavior and variations over time.
+The dataset aggregates these metrics monthly, offering understandings of traffic behavior and variations over time.
 
@@
 - Speed bins (e.g., 0-4 km/h, 5-9 km/h) for detailed breakdowns of vehicle counts across speed ranges.
-- @tbl-speed-preview showcases a simplified preview of the dataset with selected columns to highlight key features.
+
+@tbl-speed-preview showcases a simplified preview of the dataset with selected columns to highlight key features.
 
@@
 #| label: tbl-speed-preview
-#| tbl-cap: Preview of analysis traffic speed dataset.
+#| tbl-cap: Simplified preview of analysis traffic speed dataset.
 #| echo: false
@@
 
-@tbl-random-sample provides a random sample of 10 observations, offering an unbiased snapshot of key metrics such as speed percentiles, total traffic volume, and counts of vehicles exceeding 100 km/h.
+@tbl-random-sample provides a simplified random sample of 10 observations, offering an unbiased snapshot of key metrics such as speed percentiles, total traffic volume, and counts of vehicles exceeding 100 km/h.
 This selection illustrates the dataset's diversity without emphasizing specific patterns.
@@
 #| label: tbl-random-sample
-#| tbl-cap: Random sample of traffic speed dataset.
+#| tbl-cap: Simplified random sample of analysis traffic speed dataset.
 #| echo: false
@@
 
-Vehicles traveling at extreme speeds (over 100 km/h) represent a critical factor in understanding traffic safety and violations.
+Vehicles traveling at extreme speeds (over 100 km/h) represent a significant factor in understanding traffic safety and violations.
 @tbl-extreme-speeds highlights the top 5 months with the highest counts of vehicles exceeding 100 km/h.
-This provides insights into when extreme speeds are most prevalent.
+This provides understandings of when extreme speeds are most prevalent.
 
@@
 
-Another key insight from this dataset is the proportion of vehicles traveling within a moderate speed range (50–70 km/h).
+Another perspective from this dataset is the proportion of vehicles traveling within a moderate speed range (50–70 km/h).
 This measure helps understand how typical driving speeds align with safe and efficient traffic flow.
@@
 #| message: false
+
 # Summarize Key Percentiles and Calculate Speed Spread
@@
 The above model was chosen to investigate the relationship between the average speed (`spd_mean`) of vehicles and key predictors, including total traffic volume (`volume`) and speed percentiles (`pct_05`, `pct_50`, and `pct_95`).
-These variables were selected because they provide a comprehensive view of traffic patterns and vehicle behavior across different speed thresholds.
+These variables were selected because they provide a complete view of traffic patterns and vehicle behavior across different speed thresholds.
 
-The percentiles represent critical thresholds of speed distribution, with the 5th percentile (`pct_05`) capturing the slowest vehicles, the 50th percentile (`pct_50`) representing the median speed, and the 95th percentile (`pct_95`) reflecting the fastest vehicles.
+The percentiles represent significant thresholds of speed distribution, with the 5th percentile (`pct_05`) capturing the slowest vehicles, the 50th percentile (`pct_50`) representing the median speed, and the 95th percentile (`pct_95`) reflecting the fastest vehicles.
 These thresholds are expected to influence the average speed significantly.
@@
 
-The Bayesian linear regression model results, summarized in @tbl-modelsummary, provide insight into the relationship between mean traffic speed (`spd_mean`) and key predictors such as speed percentiles and traffic volume.
+The Bayesian linear regression model results, summarized in @tbl-modelsummary, provide understandings of the relationship between mean traffic speed (`spd_mean`) and key predictors such as speed percentiles and traffic volume.
 
@@
 #| message: false
-#| tbl-pos: H
+#| fig-pos: H
 
-modelsummary(mean_speed_model)
+modelsummary(list("Mean Speed Prediction" = mean_speed_model))
 ```
@@
 These results highlight the dominant role of speed percentiles in determining mean speed while raising questions about the negligible impact of traffic volume.
-Further exploration into potential interactions or omitted variables could clarify this finding.
+Further examination into potential interactions or omitted variables could clarify this finding.
 
@@
 
-While the Bayesian linear regression model employed in this paper provides meaningful insights into the relationship between key predictors and mean vehicle speed, several limitations warrant discussion.
+While the Bayesian linear regression model employed in this paper provides meaningful interpretations of the relationship between key predictors and mean vehicle speed, several limitations warrant discussion.
 As highlighted in @sec-model, the model does not account for potential interactions among predictors, such as the interplay between traffic volume and percentile speeds.
@@
 Future modeling efforts could incorporate interaction terms and hierarchical structures to better capture the complexity of traffic dynamics.
-Similarly, integrating data from diverse monitoring systems (e.g., highway cameras or GPS datasets) could provide a more comprehensive view of Toronto’s traffic behavior across different contexts.
+Similarly, integrating data from diverse monitoring systems (e.g., highway cameras or GPS datasets) could provide a more complete view of Toronto’s traffic behavior across different contexts.
 
@@
 Moreover, the cultural perspective within traffic policing, which often deprioritizes ticketing for non-criminal violations, exacerbates these issues.
-However, these challenges also present opportunities for leveraging technology to complement traditional enforcement and create a more comprehensive traffic management framework.
+However, these challenges also present opportunities for leveraging technology to complement traditional enforcement and create a more complete traffic management framework.
 
@@
 The findings in this study offer a foundation for improving traffic safety and efficiency in Toronto.
-A central takeaway is the value of automated enforcement in addressing critical challenges, including extreme speeds, variability in traffic patterns, and enforcement gaps.
+A central takeaway is the value of automated enforcement in addressing significant challenges, including extreme speeds, variability in traffic patterns, and enforcement gaps.
 Automated speed enforcement (ASE) cameras have already demonstrated significant success in Toronto, reducing both the prevalence of speeding and overall vehicle speeds in monitored areas.
@@
 
-Urban road design remains another critical area for intervention.
+Urban road design remains another significant area for intervention.
 Consistent median speeds and limited speed variability are essential for maintaining safe and efficient traffic flow.
@@
 The observational data analyzed in this study stems from the City of Toronto’s Safety Zone Watch Your Speed Program (WYSP), where radar-equipped speed signs monitor and aggregate vehicle speeds.
-While this data provides valuable insights, it also introduces limitations such as restricted spatial and temporal coverage, potential behavioral biases due to the visibility of speed signs, and an inability to capture all traffic volume accurately.
+While this data provides useful understandings, it also introduces limitations such as restricted spatial and temporal coverage, potential behavioral biases due to the visibility of speed signs, and an inability to capture all traffic volume accurately.
 This section proposes an idealized methodology to address these limitations and enhance the robustness of traffic speed data collection.
@@
 weekends), and seasons.
-This approach ensures a comprehensive understanding of traffic patterns and their fluctuations.
+This approach ensures a complete understanding of traffic patterns and their fluctuations.
 
 - Technological Enhancements
-  - Advanced technologies such as LiDAR and AI-based sensors can supplement traditional radar devices to increase measurement accuracy.
+  - Technologies such as LiDAR and AI-based sensors can supplement traditional radar devices to increase measurement accuracy.
 These technologies can detect multiple vehicles simultaneously, reducing blind spots and capturing speeds across a wider field of view.
@@
 
-- Cost Implications: Deploying advanced devices and rotational monitoring systems requires significant investment in equipment and personnel.
+- Cost Implications: Deploying devices and rotational monitoring systems requires significant investment in equipment and personnel.
 - Coordination with Stakeholders: Accessing certain road segments for device installation may require coordination with local authorities, businesses, and residents.
-- Data Management: Large-scale data collection introduces challenges related to storage, integration, and processing. Ensuring data integrity and consistency is critical for effective analysis.
+- Data Management: Large-scale data collection introduces challenges related to storage, integration, and processing. Ensuring data integrity and consistency is significant for effective analysis.
 
 Despite these challenges, the proposed framework offers a robust and realistic solution for improving traffic data collection.
-By addressing the limitations of the existing WYSP dataset, this methodology provides a foundation for more accurate and actionable insights into urban mobility.
+By addressing the limitations of the existing WYSP dataset, this methodology provides a foundation for more accurate and actionable understandings of urban mobility.
 
@@
 
-An evaluation of three years of data from Toronto's Automated Speed Enforcement (ASE) cameras provides valuable insights into the potential of automated systems to influence driver behavior and enhance traffic safety.
+An evaluation of three years of data from Toronto's Automated Speed Enforcement (ASE) cameras provides useful understandings of the potential of automated systems to influence driver behavior and enhance traffic safety.
 Conducted by researchers from SickKids Hospital and Toronto Metropolitan University, this study highlights the effectiveness of ASE in reducing speeding across multiple speed zones [@Rodrigues_2023].
@@
 - Rotational Deployment: Cameras were rotated across locations to broaden coverage and minimize behavioral biases from static monitoring.
-- Comprehensive Data Collection: Despite some limitations imposed by the COVID-19 pandemic, the study captured substantial data during periods of normal activity, enhancing its reliability.
+- Extensive Data Collection: Despite some limitations imposed by the COVID-19 pandemic, the study captured substantial data during periods of normal activity, enhancing its reliability.
 - Focused Interventions: Placement of ASE cameras in high-risk areas, such as school zones, demonstrated a targeted approach to improving road safety.
@@
 
-The WYSP dataset offers valuable observational insights into Toronto’s traffic patterns, but it is not without its limitations.
+The WYSP dataset offers useful observational understandings into Toronto’s traffic patterns, but it is not without its limitations.
 
@@
 - Focus on Safety Zones: Data collection in designated Safety Zones highlights areas with specific traffic and safety concerns.
-- Detailed Metrics: The dataset includes percentile-based speed metrics (e.g., 5th, 50th, and 95th percentiles), which are critical for analyzing traffic behavior.
+- Detailed Metrics: The dataset includes percentile-based speed metrics (e.g., 5th, 50th, and 95th percentiles), which are significant for analyzing traffic behavior.
 

### 2024-12-03 paper/paper.qmd
polish paper

@@
 title: "Analyzing Toronto Traffic Speeds: Understanding Urban Mobility, Safety, and Speed Dynamics"
-subtitle: "Exploring Traffic Patterns and Relationships Between Speed Percentiles and Volume"
+subtitle: "Examining Traffic Patterns and Relationships Between Speed Percentiles and Volume"
 author: 
